Multi-task learning in Computer Vision

### 2020-05-30 2-Background.tex
Sync
Add some fixes to all the things.

\chapter{Neural networks}
Background stuff 
\section{Fully Connected Networks}
When needed example of a transfer learning head?, when good/bad
\section{Convolutional Neural Networks}
Show some basic CNN things maybe and why needed over fully connected bois :) 
\subsection{Neural image processing}
How it different from classical features and how/why it works
\section{RNN}
Depending on what may be used

>>>>

\chapter{Neural networks}
The topics presented in this thesis are based on neural networks, especially convolutional neural networks, which are a distinct neural network variation for image processing using filters.
Here we provide a very brief high-level overview of some of the concepts that are important for the later chapters, but a more detailed overview of deep learning and neural networks can be found, for example, in \citep{deep-learning}.

\section{Convolutional Neural Networks}
The most significant component of Convolutional Neural Networks (CNNs) is the convolutional layers that are their primary building block.
Using convolutional filters allows neural networks to extract useful information from images to accomplish various tasks.
The convolutional blocks are based on the mathematical convolution operation of capturing the signal of pre-determined window size.
A CNN classifier often consists of multiple convolutional layers with other layers in between.
Successive convolutional layers build on top of one another, and the deeper layers get high activations on more specific features.
In many final models, the earlier layers are thus somewhat similar and re-usable as they often detect relatively primitive shapes, like diagonal or horizontal lines.
In this thesis, we often talk about the parameters or the weights of a network.
The number of parameters in a network tells how complex it is and how expensive it is to store.
The floating-point numbers in the various layers are these weights, which are generally 32-bit numbers, but their precision can vary.

\subsection{Parameters}
To get the total number of parameters required for a single convolutional layer, we need to determine the input size, convolution window size, and the number of filters.
For example, if the input is a 256x256 RGB image, the input size is 256x256x3.
Now, if we apply 16 filters with a window size of 5, each of the filters requires 5x5x3 parameters, bringing the total to 1200 parameters.
A fully connected layer with 1000 hidden units would require a matrix with around 200 million parameters, so it would not really be possible even to apply multiple of these layers while maintaining the image size.
The convolution allows for keeping the dimensions between successive layers and only modifying the number of channels between the layers.
Often convolutional models reduce the first two dimensions, and the number of channels increases in the deeper layers.
For this, the convolutions can use a stride parameter to skip some positions for the convolution.
The pooling layer is another way to reduce the size.
Generally used pooling layers are average and maximum pool layers. 
They are applied similarly to a convolution but capture the average or the maximum of a window.
Neither of these approaches requires any extra learnable weights.

\subsection{Embedding}
Generally, the role of the convolutional layers in a CNN model is to act as a feature extractor.
As can be seen in the number of parameters of the fully connected layer on a full-sized image, the image needs to have a smaller dense representation in order to connect a linear layer to it.
This representation can be generated with the convolutional layers.
The final representation that can be pushed into a linear classifier for some final inference is called an image embedding.
Embeddings are representations of things in relatively small continuous numbered vectors, here we consider embeddings of images, but embeddings are used, for example, for words and sentences as well.
Besides just doing classification, the embeddings can represent things, for example, faces or objects.
When a general embedding is generated it can be used to for example to evaluate if someone is the same person as the embedding or to find similar images.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/vggnet.png}
\caption{Image of an example of a convolutional neural network architecture with backbone (a), image classifier (b) and task head (c) labeled. The architecture in this image is the VGGNet \citep{VGG}.}
\end{figure}

\subsection{Backbone}
The feature extraction part of a convolutional classifier is called the backbone.
The backbone is annotated in Figure 2.1.
Finding better features is often correlated with more parameters and a larger model, but various architectural decisions can also help.
The largest convolutional models have thousands of layers.
Best models often use as large backbones as is possible within the constraints of the specific deployment.
The backbone is thus responsible for creating the image embedding that is used to produce an output.
Some tasks require feature maps taken from the middle of the backbone instead of only relying on the final layer.

\subsection{Head}
The final layers of a classifier are called the head.
The head is responsible for producing the output, and its exact architecture varies from task to task.
It could be just a single linear layer connecting the image embedding to final class probability distribution or an image segmentation model classifying each pixel in the image.
The head can thus contain whatever layers are needed to get the output, combining batch normalization, dropout, convolutional, recurrent, and linear layers depending on what is needed to solve the task.
An example visualization of how the different parts of a convolutional model are connected is shown in Figure 2.1.
The architecture would vary on a task-by-task basis, and the figure shows a basic image classifier architecture.

### 2020-05-30 3-Image-classification.tex
Add experiments and other cool stuff :)
Add some fixes to all the things.

\chapter{Image classification}
Image classification is one of the essential modern computer vision problems, where the goal is to create a model that can classify an input image into one of a set of pre-defined classes. Before the popularization of applying large Convolutional Neural Networks for this task, the most successful way of solving the problem was to use some algorithm for finding feature descriptors in a set of images to construct a Visual Bag of Words.  A linear classifier, like a Support Vector Machine, would then do the classification using the Bag of Words representation of an image. These days nearly all approaches are based on using deep CNNs, and working CNNs were deployed already in 1998 on character recognition in the form of LeNet \citep{leNet}.

@@

ImageNet \citep{imagenet} has been one of the most significant data sets for image classification, producing several significant improvements to modern computer vision techniques.
The ImageNet challenge \citep{ILSVRC} is a competition consisting of a selected collection of 1000 classes and a total of 1.2 million training images and 150 thousand test images from the ImageNet data set.
Winning models from the ImageNet competition have provided some of the essential architectures modern computer vision algorithms use.
In 2012 the winning model, AlexNet \citep{alexNet}, showed that it was possible to train deep CNNs efficiently using GPUs. Since 2012 all top-performing models showed some new improvements on how to create the most performant network architecture, for example, VGGNet from the year 2014 and ResNet \citep{resNet} from 2015, both of which have been popular models to use for Transfer Learning since.

Human accuracy on the ImageNet challenge is about 5.1\% \citep{imageNet_summary}, and ResNet achieves a top-5 error rate of 3.57\% \citep{resNet} and newer architectures even lower, but this still does not mean that image classification is a solved problem. The human performance experiment found that many of the human errors are caused by not having expert information in, for example, identifying animal species or not even being aware of the existence of a class \citep{imageNet_summary}. ObjectNet \citep{objectNet} is a dataset designed to test image classifiers with a focus on their generalizability. It contains many classes that also exist in the ImageNet dataset. However, they are in unexpected locations or have an unexpected pose, causing the high accuracy image classifiers trained on ImageNet to experience a 40-45\% accuracy drop when evaluating them on the ObjectNet images of classes shared between ImageNet and ObjectNet. This kind of adjustment is relatively easy for a human, and it shows that while the classifiers are good, they are by no means perfect.

@@

To give a formal definition of transfer learning, we will follow the definitions provided in \citep{transferSurvey2010}.
Let $\mathcal{D}$ be a domain, which consists of a feature space $\mathcal{X}$ and a marginal probability distribution P($\mathcal{X}$).
For a given domain,a task T \[\mathcal{T} = \{\mathcal{Y}, f\mathcal(X)\}\] \noindent consists of a label space $\mathcal{Y}$ for the inputs and of a predictive function $f(x)$ which produces predictions for all pairs \{$x_i, y_i$\} where $x_i \in \mathcal{X}$ and $y_i \in \mathcal{Y}$.
Given a source domain $\mathcal{D}_S$, a source task $\mathcal{T}_S$, a target domain $\mathcal{D}_T$ and a target task $\mathcal{T}_T$, where target and source are disjoint, transfer learning tries to improve the performance of $f_T(x)$ using $\mathcal{D}_S$ and $\mathcal{T}_S$.

@@

Picking which classifier to use as a base is often problem-dependent. As it is not possible to declare one network structure to be the best at all tasks, picking the best model to start with usually requires the user to compare different architectures and weighing the requirements for the problem at hand. Often though, the larger models will perform better, and there exists a correlation between performing well on ImageNet and being a good transfer learning model \citep{betterTransfer}. Though as can be seen in figure 3.1, better performance often comes at the cost of many more parameters, requiring more memory to train the model. Though just the number of parameters is not the only thing to compare as the throughput of a Resnet50 turns out to be about three times as large as the throughput of an EfficientNet-B4 even though they have a similar amount of parameters \citep{classifierPerformance}. So in the case of time-constrained inference environments, also the time required for an inference has to be taken into account when picking models.

If there is enough data, it turns out that using a pre-trained network does not provide any benefits in terms of the converged model accuracy, but it is not detrimental to performance either \citep{rethinkTransfer}. When training sufficiently long on a sufficient amount of data, the pre-trained and randomly initialized networks converged to similar accuracies but required significantly different amounts of training resources. Still, this does not mean that pre-training is useless by any means as the saved resources and getting models to converge faster are essential factors for progress, and of course, in many cases, training from scratch will not provide satisfying results.

\section{ResNet}
ResNet is one of the first effective and very deep Convolutional Neural Network architecture that was presented in 2015 and won the ImageNet challenge. Prior to the publication of ResNet, the most powerful networks were relatively shallow, like VGGNet, which has only 19 layers. One of the big issues relating to training deep networks is vanishing gradients, where gradients disappear when they are backpropagated through many layers \citep{wideResNet}. ResNets utilize residual connections around bottleneck building blocks, which allow for the networks to contain many more layers than those without them, the largest network presented in the original ResNet paper was 152 layers, totaling for around 8x increase in the number of layers when compared to earlier networks \citep{resNet}.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{imgs/resnet-block.png}
    \caption{Resnet building block \citep{resNet}}
\end{figure}

The ResNet block architecture allows for the blocks to learn the identity function more efficiently by trying to learn the residual function instead of the direct mapping. Normally a mapping is learned between ${x}$ and ${y}$ using a function ${H(x)}$ but by the same token, we can learn a residual transformation, \[{F(x) = H(x) - x}\] \noindent where ${H}$ is the mapping of two or more network layers. Both of these approaches should approximate the same functions; the difference is in how easy it is for the network to learn. Learning a transformation of ${F(x) = 0}$ would intuitively seem easier for a neural network than learning ${F(x) = x}$. As can be seen from the success of the ResNets compared to non-residual networks, this is what allows for creating very deep networks. If the transformation changes the input size, a matrix ${W}$ is necessary to map the input to the same dimensions, generating the final formula for a residual block. \[y = F(x, \{W_i\}) + W_s x\]

@@

\section{Improving model performance}
Improving model performance is not easy, but using the various types of skip-connections, such as the ones used in ResNet residual blocks, it is possible to increase the size of the network to massive sizes. A 557 million parameter model called GPipe \citep{gPipe} takes the model scaling to the extreme and requires some unique parallelism libraries to train the model. It is still only slightly better than previous models, showing that size is not the only thing that matters.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{imgs/scaling-networks.png}
    \caption{Various ways of scaling network architectures \citep{efficientNet}}
\end{figure}

@@

The other approach to finding a better architecture is to use different kinds of blocks that are better. One way of finding better building blocks is by manual experimentation. The other, like in the case of MnasNet \citep{mnas}, is to pose the problem as an optimization problem and to use Neural Architecture Search \citep{neuralSearch} to find the optimal solution. In the case of MnasNet, the goal is to maximize m in \[ACC(m) \times [\frac{LAT(m)}{T}]^w \] \noindent where ${T}$ is the target, by searching in a constrained space. So the goal is to optimize the accuracy with a given latency on a specific device. The search produces small but very efficient models when the target device is a mobile phone.

\section{EfficientNet}
EfficientNet models form a family of models ranging from EfficientNet0 to EfficientNet7 that were generated by smartly scaling existing convolutional models to optimize them for efficiency in a similar way to MnasNet \citep{efficientNet}.
The more complex models are created by using compound scaling on the EfficientNet0 model, shown in Figure 3.3 (e),  where the width, depth, and the resolution of the network are scaled using a factor ${\phi}$.
As the cost of neural architecture search grows exponentially as the size of the network grows, the search is only done on the base model.
The search of ${\phi}$ is a constrained optimization problem, where width, depth and, resolution are constrained such that ${d^\phi \times w^\phi \times r^\phi \approx 2}$ and each increase in ${\phi}$ increases network FLOPS by ${2^\phi}$.
The architecture search for the base network is similar to that of MnasNet. The target of EfficinetNet is  \[ACC(m) \times [\frac{FLOPS(m)}{T}]^w\] \noindent so the goal is to optimize for both accuracy and number of floating-point operations instead of latency as in MnasNet \citep{efficientNet}.

@@

>>>>

\chapter{Image classification}
Image classification is one of the essential modern computer vision problems, where the goal is to create a model that can classify an input image into one of a set of pre-defined classes. Before the popularization of applying large Convolutional Neural Networks for this task, the most successful way of solving the problem was to use some algorithm for finding feature descriptors in a set of images to construct a Visual Bag of Words \citep{vbow}.  A linear classifier, like a Support Vector Machine, would then do the classification using the Bag of Words representation of an image. These days nearly all approaches are based on using deep CNNs, and working CNNs were deployed already in 1998 on character recognition in the form of LeNet \citep{leNet}.
In this chapter, we will present what ImageNet \citep{imagenet} is and how the models trained with it are important when applying transfer learning, and we will show an example of how it can be done.

@@

ImageNet \citep{imagenet} has been one of the most significant data sets for image classification. 
It is has enabled the development and served as a benchmark for several significant improvements to modern computer vision techniques.
The ImageNet challenge \citep{ILSVRC} is a competition consisting of a selected collection of 1000 classes and a total of 1.2 million training images and 150 thousand test images from the ImageNet data set.
Winning models from the ImageNet competition have provided some of the essential architectures modern computer vision algorithms use.
In 2012 the winning model, AlexNet \citep{alexNet}, showed that it was possible to train deep CNNs efficiently using GPUs. Since 2012 all top-performing models showed some new improvements on how to create the most performant network architecture, for example, VGGNet \citep{VGG} from the year 2014 and ResNet \citep{resNet} from 2015, both of which have been popular models to use for Transfer Learning since.

Human accuracy in image classification on the ImageNet challenge is about 5.1\% \citep{imageNet_summary}, ResNet achieves a top-5 error rate of 3.57\% \citep{resNet} and newer architectures even lower, but this still does not mean that image classification is a solved problem. The human performance experiment found that many of the human errors are caused by not having expert information in, for example, identifying animal species or not even being aware of the existence of a class \citep{imageNet_summary}. ObjectNet \citep{objectNet} is a dataset designed to test image classifiers with a focus on their generalizability. It contains many classes that also exist in the ImageNet dataset. However, the objects in the images are in unexpected locations or have an unexpected pose, causing the high accuracy image classifiers trained on ImageNet to experience a 40-45\% accuracy drop when evaluating them on the ObjectNet images of classes shared between ImageNet and ObjectNet. This kind of adjustment is relatively easy for a human, and it shows that while the classifiers are good, they are by no means perfect.

@@

To give a formal definition of transfer learning, we will follow the definitions provided in \citep{transferSurvey2010}. 
Let $\mathcal{D}$ be a domain, which consists of a feature space $\mathcal{X}$ and a marginal probability distribution P($\mathcal{X}$). 
For a given domain, a task $\mathcal{T}$ \[\mathcal{T} = \{\mathcal{Y}, f\mathcal(X)\}\] \noindent consists of a label space $\mathcal{Y}$ for the inputs and of a predictive function $f(x)$ which produces predictions for all pairs \{$x_i, y_i$\} where $x_i \in \mathcal{X}$ and $y_i \in \mathcal{Y}$. 
Given a source domain $\mathcal{D}_S$, a source task $\mathcal{T}_S$, a target domain $\mathcal{D}_T$ and a target task $\mathcal{T}_T$, where target and source are disjoint, transfer learning tries to improve the performance of $f_T(x)$ using $\mathcal{D}_S$ and $\mathcal{T}_S$.

@@

Picking which classifier to use as a base is often problem-dependent. As it is not possible to declare one network structure to be the best at all tasks, picking the best model to start with usually requires the user to compare different architectures and weighing the requirements for the problem at hand. Often though, the larger models will perform better, and there exists a correlation between performing well on ImageNet and being a good transfer learning model \citep{betterTransfer}. Though as can be seen in figure 3.1, better performance often comes at the cost of many more parameters, requiring more memory to train the model. Though, as the results in \citep{classifierPerformance} show, just the number of parameters is not the only thing to compare as the throughput of a Resnet50 turns out to be about three times as large as the throughput of an EfficientNet-B4 \citep{efficientNet} even though they have a similar amount of parameters. So in the case of time-constrained inference environments, also the time required for an inference has to be taken into account when picking models.
Often though, the number of parameters is a relatively good proxy for the inference time.
The throughput is likely due to the differences in the internal implementations of the EfficientNet and normal CNNs.

If there is enough data, it turns out that using a pre-trained network does not provide any benefits in terms of the converged model accuracy, but it is not detrimental to performance either \citep{rethinkTransfer}. When training sufficiently long on a sufficient amount of data, the pre-trained and randomly initialized networks converged to similar accuracies but required significantly different amounts of training resources. Still, this does not mean that pre-training is useless by any means as the saved resources and getting models to converge faster are essential factors for progress. And of course, there is not enough data in many cases, and training from scratch will not provide satisfying results due to the large number of parameters that would need to be optimized.
In these cases, it would then be a decision to apply transfer learning or to get a model that is not good enough to produce reliable results.

\section{ResNet}
ResNet is one of the first effective and very deep Convolutional Neural Network architecture that was presented in 2015 and won the ImageNet challenge. Prior to the publication of ResNet, the most powerful networks were relatively shallow, like VGGNet \citep{VGG}, which has only 19 layers. One of the big issues relating to training deep networks is vanishing gradients, where gradients disappear when they are backpropagated through many layers \citep{wideResNet}. ResNets utilize residual connections around bottleneck building blocks, which allow for the networks to contain many more layers than those without them, the largest network presented in the original ResNet paper was 152 layers, totaling for around 8x increase in the number of layers when compared to earlier networks \citep{resNet}.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/resnet-block-own.png}
\caption{Resnet building block for learning the residual \citep{resNet}}
\end{figure}

The ResNet block architecture allows for the blocks to learn the identity function more efficiently by trying to learn the residual function instead of the direct mapping. The residual can be learned by adding the unchanged weights of an earlier layer to the output of some transformations on it.
The residual block is shown in Figure 3.2.
Normally a neural network layer learns the mapping between ${x}$ and ${y}$ using a function ${H(x)}$, but by the same token, we can learn a residual transformation.
Learning these residuals using the identity mapping skip-connections is the revolutionary idea introduced by the ResNet.
Since these residual connections are only connections, they do not require extra parameters to be learned and are free in terms of learnable parameters.
The residual transformation is
\[{F(x) = H(x) - x}\] \noindent where ${H}$ is the mapping of two or more network layers. Both of these approaches should approximate the same functions; the difference is in how easy it is for the network to learn. Learning a transformation of ${F(x) = 0}$ would intuitively seem easier for a neural network than learning ${F(x) = x}$. As can be seen from the success of the ResNets compared to non-residual networks, this is what allows for creating very deep networks. If the transformation changes the input size, a matrix ${W}$ is necessary to map the input to the same dimensions, generating the final formula for a residual block. \[y = F(x, \{W_i\}) + W_s x\] \noindent where $W_i$ and $W_s$ are the the weight matrices to be applied to x.

@@

\section{Improving model performance}
Improving model performance is not easy, but using the various types of skip-connections, such as the ones used in ResNet residual blocks with the identity transform, it is possible to increase the size of the network to massive sizes. A 557 million parameter model called GPipe \citep{gPipe} takes the model scaling to the extreme and requires some unique parallelism libraries to train the model. It is still only slightly better than previous models, showing that size is not the only thing that matters.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/scaling-networks-own.png}
\caption{Various ways of scaling network architectures \citep{efficientNet}. a) Shows the baseline network. b) Shows how to scale by the number of channels. c) Shows how to scale the number of layers. d) Shows how to scale by resolution. e) Shows how to scale all the metrics with a compound factor.}
\end{figure}

@@

The other approach to finding a better architecture is to use different kinds of blocks that are better. One way of finding better building blocks is by manual experimentation. The other, like in the case of MnasNet \citep{mnas}, is to pose the problem as an optimization problem and to use Neural Architecture Search \citep{neuralSearch} to find the optimal solution. In the case of MnasNet, the goal is to maximize the model architecture m in \[ACC(m) \times [\frac{LAT(m)}{T}]^w \] \noindent where ${T}$ is the target, by searching in a constrained space. So the goal is to optimize the accuracy (ACC) of the model with a given latency (LAT) on a specific device. The search produces small but very efficient models when the target device is a mobile phone.

\section{EfficientNet}
EfficientNet models form a family of models ranging from EfficientNet0 to EfficientNet7 that were generated by smartly scaling existing convolutional models to optimize them for efficiency in a similar way to MnasNet \citep{efficientNet}. 
The more complex models are created by using compound scaling on the EfficientNet0 model, shown in Figure 3.3 (e),  where the width, depth, and the resolution of the network are scaled using a factor ${\phi}$. 
As the cost of neural architecture search grows exponentially as the size of the network grows, the search is only done on the base model. 
The search of ${\phi}$ is a constrained optimization problem, where width, depth and, resolution are constrained such that ${d^\phi \times w^\phi \times r^\phi \approx 2}$ and each increase in ${\phi}$ increases network FLOPS by ${2^\phi}$. 
The architecture search for the base network is similar to that of MnasNet. The target of EfficinetNet is to maximize the model m  \[ACC(m) \times [\frac{FLOPS(m)}{T}]^w\] \noindent so the goal is to optimize for both accuracy (ACC) and number of floating-point operations (FLOPS) instead of latency as in MnasNet \citep{efficientNet}.

@@

\section{Training image classifiers}
An image classifier's basic function is to predict which of some set of classes exists in an image.
So, to train a model to answer such a question, a data set with images and accompanying labels are needed.
For example, the Oxford-IIT Pets dataset \citep{catsdogs} consists of images of 37 different cat and dog species and labels for those images signifying which of the 37 classes an image represents.
The dataset consists of about 200 images per class, so it is relatively small and an excellent example of a situation where transfer learning is instrumental in getting good results.

After the data of the images and labels are gathered to a data loader, that outputs small batches for training, the model needs to be constructed.
For the backbone, any ImageNet classifier will do. 
For example, EfficientNet-B4 could be a good decision.
So we will initialize the backbone of our model with the weights of a pre-trained EfficientNet that was trained on the ImageNet data set by someone with access to large computational resources.
After the model has been initialized, the head has to be changed to be compatible with our new task.
By default, the model will be expecting to classify images to the thousand classes of the ImageNet.
In this case, we want to solve the 37 cat and dog species problem, so the final fully connected layer has to be changed.
It will be initialized as a fully connected layer with 37 output units, representing each of the classes, connecting the embedding to the outputs.
The head does not need to be just a single, fully connected layer, but it could also contain more fully connected layers, dropout layers, and batch normalization layers and more.
Adding more than a single fully connected layer might be a good idea if the backbone remains frozen and is not fine-tuned. 
That way, the head can learn some useful combinations of the image embedding instead of optimizing the features in the backbone.

Once all the layers are in place, the model needs to be optimized with a training loop.
Every iteration, a loss needs to be calculated for the batch that has been sampled from the data loader.
To be able to calculate the loss, the outputs of the final fully connected layer need to be normalized into a distribution of probabilities with softmax.
The softmax is defined as
\[{Softmax(x_i) = \frac{exp(x_i)}{\sum_j exp(x_j)}}\] \noindent
For each class $i$, there is an output score of $x_i$, for which we will calculate the proportion of its score with respect to the sum of all the classes' scores $x_j$, creating the probability distribution.
Since we have a single correct class, we would want to maximize the probability of the correct class and minimize the probability of the other classes.
The loss function to do exactly this optimization is called categorical cross-entropy, and it is defined as follows.
\[{CE = -\frac{1}{N}\sum_{i=1}^{N}(y_i * log(p_i) + ( 1 - y_i ) * log ( 1 - p_i))}\] \noindent
Where $N$ is the number of classes, $y_i$ is either 1 or 0 depending on if it is the correct label and $p_i$ is the probability for that class.
This formula does exactly what we previously described that we wanted to do.
To get the loss for a single class, if $y = 1$, the value will be the first part of the sum, if $y = 0$, the value will be the second part of the sum, since $(1 - y_i)$ is now 1.
When we optimize the model, in the case of the chosen architecture, the model will reach around 95\% accuracy on this data set \citep{efficientNet}.

### 2020-05-30 4-Object-detection.tex
Add some fixes to all the things.

\section{Training object detectors}

>>>>

\section{Training object detectors}
As we described previously, object detectors are trained when there is a need to localize some set of classes within images by describing them with bounding boxes.
One widely used format for labeling such a dataset is the format for the object detection data in the COCO data set.
The COCO data consists of a set of images and a separate JSON file containing the labels.
The label file contains a list of all the images for the data set and a unique identifier for each image.
There is also an annotation list that contains a list of objects that specify the bounding box annotations.
The annotations are specified as being linked to a single image by its image id and defining the bounding box annotation dimensions.
The bounding box annotations are formatted by providing the coordinates of the top-left corner of the annotations and then the width and height of the box, so it is a tuple of four numbers (x, y, width, height).
To train on new data sets, they should be formatted similarly to make the training as easy as possible.

Once the data set has been gathered and collected in a correct format for the object detection task, a model needs to be picked.
Here, the final inference model's requirements need to be evaluated to decide whether a single-stage or a two-stage detector should be used.
For example, when running the model in real-time, a single-stage detector might be better, and in the case of doing text detection, it seems to be more popular to use a two-stage detector.
We will describe here how things work for the EfficientDet, which is a single-stage detector.

Once the exact architecture has been chosen, an implementation of it is needed to construct the model.
Generally, all the popular architectures have open-source implementations on GitHub in both Tensorflow and Pytorch, which are the most popular deep learning frameworks, that can be used.
Most of these implementations come quite readily wrapped in a way that given a correctly formatted data set, the models will be automatically initialized with the correct number of classes, and the training process is already specified by the input of a few necessary parameters.
As mentioned in the case of image classification, it is extremely important to initialize with the pre-trained weights with an object detection model to achieve the best results instead of training from scratch.
In an object detector, the problem-specific part is the number of classes in the classification head, which can be seen in Figure 5.3.
For the rest of the network, pre-trained weights should be used.

Training the EfficientNet itself is a multi-task training problem. 
We have two losses, the regression loss signifying how close the bounding boxes are to the annotations by calculating IoU and classification loss, which tells how incorrect the predicted labels are.
A major problem in optimizing anchor-based single-stage detectors like the EfficientDet is the fact that between the background no-class and the foreground classes of interest, there is a massive discrepancy in the number of instances to classify \citep{retinaNet}.
The anchor-based detectors determine whether an anchor placed at any valid spot in the image contains a class of interest or the background. 
There are tens of thousands to hundreds of thousands of valid anchors to consider depending on the architecture. 
Due to this, we can have a class imbalance of 1:1000 between the positive and negative cases.
For this reason, the recent single-stage detectors use a focal loss presented in the RetinaNet paper, to account for this class imbalance \citep{retinaNet}.

The focal loss introduces two new parameters to the cross-entropy formula to account for the class imbalance. 
The formula for focal loss (FL) is as follows.
\[FL(p_i) = -\alpha_i * log(1-p_i)^\gamma * log(p_i)\] \noindent
Where $\alpha$ and $\gamma$ are the new scaling factors and $p_i$ is defined as $p$ if y = 1 and $1 - p$ otherwise.
The factor $\alpha$ is there to weight the loss for the positive and negative cases separately. 
It is a parameter that is recommended to be set at the inverse class frequency, but the value can be searched by experimentation as well \citep{retinaNet}.
The factor $\gamma$ makes sure that a large number of easy classifications do not dominate the total loss.
For $\gamma$ the recommended value is two \citep{retinaNet}.
With this loss function, it is possible to have a very large number of anchors, where a large majority of them are always predicted as the background class.

With these, the object detection model can be trained.
Still, just leaving all the extra parameters at the default values may not produce the best results.
For example, the possible anchors that the model uses are specified using specific parameters.
The parameters that can be modified are the anchor size, depending on whether large or small objects need to be detected, the anchor sizes can be modified to suit the task at hand.
The anchors also have aspect ratios applied to them, which can be modified, the default ratios are 1:1, 2:1, and 1:2.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{imgs/nms.png}
    \caption{Object detection without non-max suppression. The same car is detected multiple times.}
\end{figure}

Finally, once a model has been trained with a suitable dataset, architecture, and hyperparameters, it can output predictions for new images.
If the model is used without any modifications, the output boxes will likely look like the predictions in Figure 5.4.
To fix this issue in the predictions, non-max suppression (NMS) should be used \citep{nms}.
With non-max suppression, we can achieve only a single bounding box that better represents what is most likely wanted.
In some cases, using NMS can lead to not being able to predict heavily overlapping classes.
Generally, though NMS should pretty much always be used to avoid the issue depicted in the Figure 5.4.

The object detectors are a good example of multi-task learning and how using the same feature maps, it is possible to solve multiple problems.
In this case, only a single annotation is needed to label for the two different tasks that can be optimized together.
Since we have two losses that are calculated, we can decide to focus on one of the losses by adding a multiplier to it and see if the higher loss could improve that task.
While some of the things covered here will be similar in the two-stage detectors, the region proposal network approach is very different from the anchor-based single-stage detectors covered in this section.
